Maenpuen et al. Appendix S1-7: Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per area (LMA)

### 2022-05-06 ms/LMA_method.Rmd
update result

@@

To quantify the effect of leaf morphology and leaf punch size on the relationship between whole-leaf and leaf disc LMA, we built a hierarchical Bayesian model.
The whole-leaf LMA was assumed to follow a normal distribution (N) with the mean of leaf disc LMA and covariates of  leaf tissue density, leaf area and leaf thickness on the log-scale;

@@

*Bivariate trait relationships for mean and CV*

Bivariate trait relationships between disc-based and whole-leaf log-transformed LMA were further summarized with variance explained (*R^2^*) based on Pearson correlations and with standardized major axis (SMA) regressions [@Warton2006] across all the samples and different leaf area and thickness groups.

Species groups were divided into four categories based on the medians of leaf area and leaf thickness across all species:
thin and small-leaved, thin and large-leaved, thick and small-leaved, thick and large-leaved species.
Median values are as following:
small-leaved species (LA < `r val$LA_mid` cm^2^),
and thin-leaved species (LT < `r val$LT_mid` mm).

@@

SMA analyses were run using the R package *smatr* [@Warton2012a]; this included tests for slope heterogeneity and elevation differences between disc-based and whole-leaf estimates of LMA values.
Confidence intervals of SMA estimates were based on 2000 bootstraps.
All statistical analyses were conducted in R version 4.1.3 [@RCoreTeam2022].

>>>>

*Relationship between whole-leaf and leaf disc LMA*

A bivariate trait relationships between disc-based and whole-leaf log-transformed LMA were summarized with variance explained (*R^2^*) based on Pearson correlations and with standardized major axis (SMA) regressions [@Warton2006] across all the species.

SMA analyses were run using the R package *smatr* [@Warton2012a]; this included tests for slope heterogeneity and elevation differences between disc-based and whole-leaf estimates of LMA values.
Confidence intervals of SMA estimates were based on 2000 bootstraps.

@@

To further quantify the effect of leaf morphology and leaf punch size on the relationship between whole-leaf and leaf disc LMA, we built a hierarchical Bayesian model.
The whole-leaf LMA was assumed to follow a normal distribution (N) with the mean of leaf disc LMA and covariates of  leaf tissue density, leaf area and leaf thickness on the log-scale;

@@

We also perfomred ordinary least-squares (OLS) regressions to see the relationship between whole-leaf to leaf disc LMA ratio and petiole to leaf dry mass ratio and between leaf area petiole to leaf dry mass ratio, using a subset of the dataset in which we measured leaf lamina and petiole dry mass separately.

*Relationship between disc-based and whole-leaf estiamtes of intraspecific variation*

@@

SMA analyses were run as described above.
All statistical analyses were conducted in R version 4.1.3 [@RCoreTeam2021].
